### 2014-11-29 presentation/pres.tex
presentation

\documentclass{beamer}

\usepackage{array}

% \usepackage{beamerthemesplit} // Activate for custom appearance

\title{Authentication Using Keystroke Timing}
\author{Noam Zilberstein and Jonathan Chen}
\date{\today}

\begin{document}

\frame{\titlepage}

\section[Outline]{}
\frame{\tableofcontents}

\section{Introduction}
\subsection{Project Motivation}
\subsection{Project Goals}
\subsection{Related Work}
\section{The Protocol}
\subsection{The Basic Protocol}
\subsection{The Zero Knowledge Protocol}
\section{Challenges and Limitations}
\section{Demo}

\frame
{
  \frametitle{Project Motivation}

  \begin{itemize}
  \item<1-> Authenticate users without password
  \item<2-> `Hardware' authentication factor without the need of hardware
  \item<3-> Login system that is easier and more secure than what is available today
  \end{itemize}
}

\frame
{
  \frametitle{Project Goals}

  \begin{itemize}
  \item<1-> Develop a prototype login system
    \begin{itemize}
    \item<2-> Authentication using keystroke dynamics
    \item<3-> Ease of use (user only needs to type their name)
    \end{itemize}
  \item<4-> Make the protocol zero-knowledge
    \begin{itemize}
    \item<5-> Keystroke timings are never transmitted; can't be compromised
    \end{itemize}
  \item<6-> Determine feasibility of a full-scale implementation
  \end{itemize}
}

\frame{
  \frametitle{Related Work}
  \begin{itemize}
    \item<1->1997 Monrose and Rubin, authentication scheme using keystroke dynamics
    \item<2->2005 Ara\'{u}jo et al, authentication scheme with $\sim99\%$ accuracy
    \item<3->2010 Stefan and Yao, keystroke dynamic authentication against synthetic forgeries
    \item<4->But is it possible to have a highly accurate system that is zero-knowledge?
  \end{itemize}
}

\frame
{
  \frametitle{Basic Protocol (Account Creation)}

  \begin{center}
\begin{tabular}{m{1.5in}cm{1.5in}}
\textbf{Prover (Client)} && \textbf{Verifier (Server)}\\
User types name ($A$) in web client. Keystroke vector ($V$) is recorded &$\xrightarrow{(A, V)}$ & $(A,V)$ is stored in the database
\end{tabular}
\end{center}
\vspace{1in}
Note: Very insecure! Everything is transmitted in the clear

}

\frame
{
  \frametitle{Basic Protocol (Authentication)}

  \begin{center}
\begin{tabular}{m{1.5in}cm{1.5in}}
\textbf{Prover (Client)} && \textbf{Verifier (Server)}\\
User types name ($A$) in web client. Keystroke vector ($\tilde V$) is recorded &$\xrightarrow{(A, \tilde V)}$ & Lookup key $A$ in the database to get $V$\\
Report result&$\xleftarrow{f(V, \tilde V)}$ & Send decision
\end{tabular}
\end{center}
\vspace{0.75in}
Where $f(V, \tilde V)\in \{0,1\}$  decides if the (noisy) timing vector $\tilde V$ is close enough to the expected timing vector $V$\\
Note: Also very insecure!

}

\frame
{
  \frametitle{Zero Knowledge Protocol (Account Creation)}

  \begin{center}
\begin{tabular}{m{1.5in}cm{1.5in}}
\textbf{Prover (Client)} && \textbf{Verifier (Server)}\\
Choose some generator $g$ & $\xrightarrow{g}$ & Remember $g$\\
User types name ($A$) in web client. Keystroke vector ($V$) is recorded. Let $h = g^{f(V)}$ &$\xrightarrow{h}$ & $(A,g,h)$ is stored in the database
\end{tabular}
\end{center}
\vspace{1in}
Data is transmitted in the clear. We don't care

}

\frame
{
  \frametitle{Zero Knowledge Protocol (Authentication)}

  \begin{center}
\begin{tabular}{m{1.5in}cm{1.5in}}
\textbf{Prover (Client)} && \textbf{Verifier (Server)}\\
User types name ($A$) in web client. Keystroke vector ($\tilde V$) is recorded &$\stackrel{A}{\longrightarrow}$ & \\
&$\stackrel{g}\longleftarrow$ & Lookup key $A$ in the database to get $g$ and $h$\\
Choose $r\in_R\mathbb{Z}_g$, let $a = g^r$ & $\stackrel{a}\longrightarrow$ & \\
& $\stackrel{b}\longleftarrow$ & Choose $b\in_R\mathbb{Z}_g$ \\
Let $c = r + f(\tilde V)\cdot b$ & $\stackrel{c}\longrightarrow$ & \\
Report decision & $\stackrel{ah^b \stackrel{?}{=} g^c}\longleftarrow$ &
\end{tabular}
\end{center}
Where $f$ is some fuzzy extractor
}

\frame
{
  \frametitle{Zero Knowledge Protocol}
  \begin{itemize}
    \item<1->Correctness: $ah^b = g^r(g^{f(V)})^b = g^{r+f(V)\cdot b} \stackrel?= g^{r+f(\tilde V)\cdot b} = g^c$
    \begin{itemize}
      \item<2-> This will hold if $f(V) = f(\tilde V)$
    \end{itemize}
    \item<3-> Security reduces to discrete log
    \item<4-> ...But can the fuzzy extractor accurately authenticate someone?
  \end{itemize}

}
\frame
{
  \frametitle{Challenges and Limitations}
  \begin{itemize}
    \item<1-> Zero knowledge means information loss  
    \item<2-> The protocol is only as good as the fuzzy extractor
    \begin{itemize}
      \item<3-> Accuracy vs Precision tradeoff
      \item<4-> User expects to be authenticated first try
      \item<5-> Precision varies by user
    \end{itemize}
    \item<6->Conclusion: performance of machine learning approach is likely not attainable with zero-knowledge
  \end{itemize}

}
\frame{\frametitle{Demo}}

\end{document}